---
title: 'Tidytuesday #2: Passwords'
author: Zach Bogart
date: '2020-08-31'
slug: tidytuesday-2-passwords
categories:
  - R
tags:
  - tidytuesday
subtitle: "Data on Cracking (Bad) Passwords"
image:
  caption: ''
  focal_point: ''
  preview_only: no
output: html_document
editor_options:
  chunk_output_type: console
links:
  koji_button:
    icon: mug-hot
    icon_pack: fas
    name: Give a Coffee
    url: https://ko-fi.com/zachbogart
---

<!--Starting Blurb-->
A quick exploration, looking at password data from [TidyTuesday](https://github.com/rfordatascience/tidytuesday).

<!--Put actual work here-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache.lazy = FALSE,
                      fig.width = 8,
                      fig.height = 5)
```

## Setup

```{r message=FALSE}
library(tidyverse)
library(tidytuesdayR)
theme_set(theme_light())

plot_caption = "zachbogart.com\nSource: tidytuesday"
```

Needed a little cleaning. The [help file](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-01-14) said the strength of a password is between one and ten, but there are some outliers. Also some `NA`s to remove.

```{r cache=TRUE, message=FALSE}
tt <- tt_load("2020-01-14")

passwords <- tt$passwords %>% 
  filter(between(strength, 1, 10)) %>% 
  filter(!is.na(.)) %>% 
  mutate(length = str_length(password))
```

## What are these categories of passwords?
Passwords are split into "categories". Most passwords are stated as a `name`. 

```{r}
passwords %>% 
  count(category, sort=TRUE)
```

## Use Longer Passwords

Looks like longer passwords take longer to crack. Makes sense.

```{r}
# what passwords take a while to crack?
passwords %>% 
  arrange(desc(offline_crack_sec))
```

Let's look at how the average time to crack a password (offline) increases as password length increases.

```{r}
# do longer passwords take longer to crack?
crack_time <- passwords %>% 
  filter(length <= 8) %>% 
  group_by(length) %>% 
  summarise(avg_offline_time = mean(offline_crack_sec),
            n = n())

crack_time %>% 
  ggplot(aes(length, avg_offline_time)) +
    geom_point(color = "#50C779") +
    geom_line(color = "#50C779") +
    scale_y_log10() +
    labs(title = "Length Matters",
         subtitle = "Average Time to Crack (Bad) Passwords vs. Password Length",
         caption = plot_caption,
         x = "Password Length",
         y = "Average Time to Crack (seconds)")
```

It's exponential! Adding another character makes it much harder (a lot more to brute force, multiplying by a factor).

## Do Numbers Matter?

Wondering if passwords with more numbers means a better password. Looks like most passwords don't have numbers. And no, using numbers doesn't mean you have a more secure password (the computer doesn't care; better off with [an XKCD-style password](https://xkcd.com/936/)).

```{r}
# do "better" passwords have numbers?
passwords %>% 
  mutate(number_count = factor(str_count(password, "[0-9]"))) %>%
  
  ggplot(aes(number_count, strength, color = number_count)) +
    geom_jitter() +
    labs(title = "Numbers Don't Help",
         subtitle = "Password Strength vs. Numbers Used in Password",
         caption = plot_caption,
         x = "How Many Numbers Used",
         y = "Password Strength",
         color = "How Many Numbers Used")
```

<!--Section on Learning (general notes that came up)-->
## Learning

- Was having trouble setting fixed color across multiple layers in a ggplot. 
- `geom_jitter` is useful for expanding overlapping continuous data, but it is probably most helpful when plotting two continuous values. Otherwise you get bands of data.
- If any of these passwords look familiar, go get a password manager!

<!--Noun Image Credit-->
#### Image Credit
[jail](https://thenounproject.com/search/?creator=4129988&q=lock&i=2585095) by Zach Bogart from [the Noun Project](https://thenounproject.com/)
